\chapter*{Acknowledgements}

I would like to start thanking to Funda\c{c}\~{a}o para a Ci\^{e}ncia e a Tecnologia (FCT) for the support to this work under the grant No.\textbf{PD/BD/114306/2016} carried out as part of the training in the framework of the Advanced Program in Plasma Science and Engineering (APPLAuSE, sponsored by FCT under grant No.~PD/00505/2012). In addition a special gratitude to EUROfusion for the provided support in several missions. 
\smallskip

I would like to thank my supervisor Prof. Horácio Fernandes for all the support and the patience he had with me through all these years, for always guiding me and for his trust on me. Thank you for always finding the time and space to work together and thank you for the  freedom to experiment and apply practically every idea born out if this research brought in ISTTOK.
\smallskip

Thanks to my co-supervisor Prof. Nuno Cruz for his unconditional support at any moment and for guiding me through this project even before I knew we were going to work together. From JT60-SA to the most specific ISTTOK details, he was always there helping me to understand.
\smallskip

I would like to thank my external co-supervisor Prof. Alfredo Pironti and the CREATE team for being a constant support during the development of the magnetic control for JT60-SA  and for welcoming me in Federico II and Naples. \smallskip

Thanks to all my colleagues and friends I met at CREATE in Naples: Adriano, Gianmaria, Alejandro and Suping. Grazzie mille.\smallskip

I would like to thank my work colleagues and friends at IPFN and ISTTOK for your constant help and for giving me your support even during the nights and weekends. Thanks to Hugo and Humberto for helping me to operate and understand ISTTOK. Thanks to Prof. Bernando Carvalho and Prof. Rui Coelho for all your time and for the knowledge you shared with me. A todos vocês muito obrigada.\smallskip


A very caring thank you to Polina and Ridhima who were always there for me and became my family during these years. May our friendship last for many years ahead and may your goals and dreams come true.\smallskip

Thank you Dori for adventuring yourself in trying to understand these concepts and helping me with some of the drawings in this work. I know more adventures are coming ahead for us both.\smallskip

A mis papás por su constante apoyo, los quiero.   

 





